---
title: "08RS data"
number-sections: true
format:
  html:
    toc: true
    toc-depth: 4
editor: source
editor_options: 
  chunk_output_type: console
#bibliography: references.bib
#csl: the-american-naturalist.csl
---

```{r include = FALSE, eval = FALSE}
# The code that adds the .nojekyll file to the GitHub repository:
file <- ".nojekyll"
file.create(file)
gert::git_add(file)
gert::git_commit("Adding the .nojekyll file")
gert::git_push()
rm(file)
```

```{r include = FALSE, eval = FALSE}
run_all <- function() {
  knitr::purl("data.qmd", "tmp.R", documentation = FALSE)
  source("tmp.R")
  file.remove("tmp.R")  
}

rerun_all <- function() {
  rm(list  = grep("run_all", ls(envir = .GlobalEnv), value = TRUE, invert = TRUE),
     envir = .GlobalEnv)
  run_all()
}
```

```{r include = FALSE}
setHook("plot.new", NULL, action = "replace")

par2 <- function(...) par(..., mgp = c(1.5, .5, 0), bty = "n")

knitr::knit_hooks$set(
  margin = function(before, options, envir) {
    if (before) par2(plt = c(.105, .97, .15, .95)) else NULL
  })

figw <- 7  # default is 7 and maximum possible is 8.3
figh <- 5  # default is 5
eps <- .7

knitr::opts_chunk$set(margin     = TRUE,
                      fig.retina = 2,
                      fig.align  = "center",
                      fig.width  = eps * figw,
                      fig.height = eps * figh) 
```


## Constants

The path to the data and cache folders on the local computer:

```{r}
if (Sys.getenv("USER") == "MarcChoisy" | Sys.getenv("USER") == "marc") {
  OneDrive <- "~/Library/CloudStorage/OneDrive-OxfordUniversityClinicalResearchUnit/"
  data_folder  <- paste0(OneDrive, "SharePoint/OUCRU Modelling/08RS/data/")
  cache_folder <- paste0(OneDrive, "GitHub/choisy/08RS/cache/")
}
if (! dir.exists(cache_folder)) dir.create(cache_folder)
```

```{r include = FALSE}
make_path <- function(x) paste0(cache_folder, x)
file_exists <- function(x) file.exists(make_path(x))
readRDScache <- function(x) readRDS(make_path(x))
saveRDScache <- function(object, file) saveRDS(object, make_path(file))
```

Saras' CSV file:

```{r}
saras_file <- "complete data including all withdrawals_updated26_3_21.csv"
```

08RS's CRF file:

```{r}
CRF08RS_file <- "6-11-2024-CTU08RS_Data.xlsx"
```

03EI's CRF file:

```{r}
CRF02EI_file <- "6-11-2024-CTU02EI_Data.xlsx"
```

PCR results data file:

```{r}
PCR_file <- "03EI-08RS PCR-Seq result.xlsx"
```

MRI data file:

```{r}
MRI_file <- "part_dataMRIentry_28AUG15_errorcor.csv"
```

Bayley's data file:

```{r}
Bayley_file <- "12-9-2025-Bayley_v3_P1_Data.xlsx"
```


## Packages

Required packages:

```{r}
required <- c("readxl", "purrr", "dplyr", "magrittr", "tidyr", "anthro", "twang",
              "cobalt", "survey", "tmle", "RColorBrewer", "lubridate", "randomForest",
              "tibble")
```

Installing those that are not installed yet:

```{r}
to_install <- required[! required %in% installed.packages()[,"Package"]]
if (length(to_install)) install.packages(to_install)
```

Loading some packages for interactive use:

```{r message = FALSE}
library(dplyr)
library(purrr)
library(stringr)
library(tidyr)
library(twang)
library(cobalt)
library(survey)
library(tmle)
library(lubridate)
library(randomForest)
library(tibble)
```


## Functions

A function that tunes a reading function:

```{r}
make_data_reader <- function(f) function(file, ...) f(paste0(data_folder, file), ...)
```

Tunings of `readxl::excel_sheets()`, `readxl::read_excel()` and `readr::read_csv()`
functions:

```{r}
list(excel_sheets2 = readxl::excel_sheets,
     read_excel2   = readxl::read_excel,
     read_csv2     = readr::read_csv) |> 
  map(make_data_reader) |> 
  list2env(envir = .GlobalEnv)
```

A function that reads all the tabs of an excel file in the data folder `data_folder`
defined above:

```{r}
read_excel_file <- function(file) {
  file |> 
    excel_sheets2() |> 
    set_names() |> 
    map(~ read_excel2(file, .x))
}
```

A function that removes some slots of a list, by names:

```{r}
remove_slots <- function(lst, slt) lst[setdiff(names(lst), slt)]
```

Below is a function that extracts some variables of some slots of a list `x` of data
frames:

```{r}
get_vars <- function(sel, x) {
  x |> 
    magrittr::extract(names(sel)) |>
    map2(sel, ~ select(.x, !!!.y))
}
```

The variables in questions are defined in the named list `sel` of character vectors.
The names of this list should be among the names of `x` and the character vectors of
each slots should be among the names of the columns of the data frames in the
corresponding slots. Here is an example where all the data frames of the list happen to
have the smes dimensions, but this does not have to be the case:

```{r}
nb_cols <- 4
nb_rows <- 3
nb_df <- 4

f_ <- function(x = 0) {
  cols <- 1:nb_cols
  cols |>
    map(~ rnorm(nb_rows)) |> 
    setNames(letters[cols + x]) |> 
    as_tibble()  
}

(x_ <- seq(0, by = nb_cols, le = nb_df) |> 
  map(f_) |> 
  setNames(paste0("df", 1:nb_df)))


(sel_ <- list(df1 = c("a", "c"), df3 = "j"))


get_vars(sel_, x_)
rm(nb_cols, nb_rows, nb_df, f_, x_, sel_)
```

Shortcut to `magrittr::extract()`:

```{r}
mextract  <- function(...) magrittr::extract(...)
```

Shortcut to `magrittr::extract2()`:

```{r}
mextract2 <- function(...) magrittr::extract2(...)
```

A function that patches data values from the data frame `patch_df` into the data frame
`into_df`, using the `key` variable as a common key between the two data frames:

```{r}
patch <- function(patch_df, into_df, key) {
  ref <- into_df[, key]
  sel <- unlist(ref) %in% patch_df[[key]]
  
  tmp <- into_df[sel, ]
  tmp <- tmp[, key] |>
    left_join(patch_df) |> 
    bind_cols(remove_slots(tmp, names(patch_df))) |> 
    mextract(names(tmp))
    
  into_df[! sel, ] |> 
    bind_rows(tmp) |> 
    left_join(x = ref, y = _, by = key)
}
```

An example:

```{r}
(into_df_ <- iris |>
  head() |> 
  as_tibble() |> 
  mutate(key_ = row_number(), .before = everything()))

(patch_df_ <- into_df_ |> 
  slice_sample(n = 4) |> 
  select(1, 3:4) |> 
  mutate(across(-1, ~ rnorm(4))))

patch(patch_df_, into_df_, "key_")
rm(patch_df_, into_df_)
```

A function that splits a data frame into a list of data frames:

```{r}
split_df <- function(x, n_rows) {
  nb_rows <- nrow(x)
  split(x, gl(nb_rows %/% n_rows + (nb_rows %% n_rows > 0), n_rows, nb_rows))
}
```

A function that appends a data frame `x` with `n` rows of values `v`:

```{r}
append_dataframe <- function(x, n = 1, v = 0) {
  1:ncol(x) |>
    map(~ rep(v, n)) |> 
    as.data.frame() |> 
    setNames(names(x)) |>  
    (\(y) bind_rows(x, y))()
}
```

A function that applies `append_dataframe()` to the last slot of a list `x` of data
frame so that the number of rows of the data frame in the last slot is equal to the
number of rows of the data frame in the first slost:

```{r}
append_last <- function(x, v = 1) {
  nb_slots <- length(x)
  nb_rows1 <- nrow(x[[1]])
  nb_rows2 <- nrow(x[[nb_slots]])
  if (nb_rows2 < nb_rows1) {
    x[[nb_slots]] <- append_dataframe(x[[nb_slots]], nb_rows1 - nb_rows2, v)
  }
  x
}
```

A tuning of the `image()` function:

```{r}
image2 <- function(x, y, z, ...) image(x, y, t(z[nrow(z):1, ]), ...)
```

A function that adds an y zero value to both ends of a data frame with two columns `x`
and `y`:

```{r}
adding_zero_ys <- function(x) {
  x <- as_tibble(x[c("x", "y")])
  x <- bind_rows(head(x, 1), x, tail(x, 1))
  x$y[c(1, nrow(x))] <- 0
  x
}
```

A function that converts a 1-row matrix with columns names into a named vector:

```{r}
as_vector <- function(x) setNames(as.vector(x), colnames(x))
```

A tuning of the `coef()` function:

```{r}
coef2 <- function(x) last(coef(x))
```

A tuning of the `confint()` function:

```{r}
confint2 <- function(x) as_vector(last(suppressMessages(confint(x))))
```

A function that retrieves the p value of the last parameter of a model:

```{r}
get_p <- function(x) last(as.vector(coefficients(summary(x))))
```

A tuning of the `svyglm()` function:

```{r}
svyglm2 <- function(formula, data, w) {
  data |>
    svydesign(ids = ~1, weights = w, data = _) |> 
    svyglm(formula, design = _)
}
```

A function that pads density coordinates with y zero value points:

```{r}
pad_density <- function(x) {
  tmp <- x$x
  x$x <- c(tmp[1], tmp, last(tmp))
  x$y <- c(0, x$y, 0)
  x
}
```

A tuning of the `density()` function:

```{r}
density2 <- function(...) pad_density(density(...))
```

A tuning of the `polygon()` function:

```{r}
polygon2 <- function(...) polygon(..., border = NA)
```

A function that computes the range of a variable `var` across a list `x` of data
frames:

```{r}
range2 <- function(x, var) {
  x |> 
    map(mextract2, var) |> 
    range()
}
```

A tuning of the `mtext()` function:

```{r}
mtext2 <- function(...) mtext(..., cex = 2 / 3)
```

A tuning of the `hist()` function:

```{r}
hist2 <- function(x, ...) hist(x, function(x) min(x):max(x), ...)
```

A tuning of the `max()` function:

```{r}
max2 <- function(x) {
  if (all(is.na(x))) return(NA)
  max(x, na.rm = TRUE)
}
```


## Reading data files

### Saras' CSV file

Reading Saras' CSV file:

```{r}
saras <- read_csv2(saras_file, col_select = -1, show_col_types = FALSE)
```

An overview of Saras' data:

```{r}
select(saras, waste, visitM, ddifENB, ddifEN, FUP, FUP1)
```

The waste variable:

```{r}
table(saras$waste)
```


### Raw data

#### 08RS CRF

Loading the data from CliRes:

```{r}
CRF08RS <- read_excel_file(CRF08RS_file)
```

The names of the data frames in CliRes and in Saras' code, with definitions:

```{r}
# CliRes        Saras           Definition
# ------------------------------------------------------------------
# ENROL         data_EN         enrollment
# HIST          data_HIST       history at enrollment
# CONHIST       CONHIST         contact history at enrollment
# EXAM          data_EX         symptoms and signs at enrollment
# LAB           data_LAB        lab results at admission
# NEU           data_NEU        neurological exam
# DAILY         data_Daily      daily review
# MED           data_MED        medications
# DEVSOCSED     data_DEV        development and socio-economic data
# DISC          data_DISC       discharge summary
# FUP           data_FUP        first follow-up day 7-10
# FUP_II        data_FUP6m      first follow-up month 6
# FUP_III       data_FUP18m     first follow-up month 18
# NEURO         data_NEURO      neurological assessment
# ABC           data_MABC       movement ABC-2
```

The 08RS CRF dictionary:

```{r}
CRF_dict <- list(
  devsocsed = list(MomEdu = c("Never been to school",
                              "Attended some primary school",
                              "Completed primary school (5th gr)",
                              "Completed lower secondary school (9th gr)",
                              "Completed higher secondary school (12th gr)",
                              "Completed university/college degree",
                              "Completed postgraduate degree"),
                   Toilet = c("Own flush toilet",
                              "Shared flush toilet",
                              "Traditional pit toilet",
                              "Ventilation improved pit toilet",
                              "No facility/bush/field",
                              "None of above"),
                   Water  = c("Private tap",
                              "Public standpipe",
                              "Bottled water",
                              "Well in own residence",
                              "Public well",
                              "Rain water",
                              "Spring",
                              "River/lake/pond", NA,
                              "None of the above")),
  disc = list(GradeHFMD   = c("grade 1",
                              "grade 2a",
                              "grade 2b(1)",
                              "grade 2b(2)",
                              "grade 3",
                              "grade 4",
                              "Not Applicable"),
              Outcome     = c("Full recovery without complication",
                              "Incomplete recovery",
                              "Transferred to another hospital",
                              "Taken home without approval",
                              "Death",
                              "Discharged to die")))
```

Selection of variables from the 08RS CRF:

```{r}
selection08RS <- list(ENROL     = c("ParNo", "DateEnrol", "Gender", "DateBirth"),
                      HIST      = c("ParNo", "DateIllness", "DateAdmHTD", "DateAdmHTD",
                                    "DateAdmHosp", "HFMDToday", "HFMDAdmitted"),
                      EXAM      = c("ParNo", "headCircumference", "height", "weigh"),
                      DEVSOCSED = c("ParNo", "MomEdu", "Toilet", "Refrigerator",
                                    "AirConditioner", "Motorbike", "Water"),
                      DISC      = c("ParNo", "DateDisc", "GradeHFMD", "TreatSepsis",
                                    "Outcome", "Seizure", "Hypertonicity", "LimbPara",
                                    "CNP", "DiapWeak", "Trache", "Nasotube",
                                    "BehaveChange"))
```


#### 02EI CRF

Reading the 02EI data file:

```{r}
CRF02EI <- read_excel_file(CRF02EI_file)
```

Selection of variables from the 02EI CRF:

```{r}
selection02EI <- list(Demo  = c("studyCode", "height", "weight"),
                      Hist  = c("studyCode", "IllnessDate"),
                      Disch = c("studyCode", "seizures", "tracheostomy",
                                "muscleStength", "limbParalysing", "nerveParalysing"))
```


#### PCR data

Reading the PCR results data:

```{r}
PCR <- PCR_file |>
  read_excel2("08RS") |> 
  select(ID, `OUCRU RESULT`) |> 
  mutate(across(ID, as.numeric),
         across(`OUCRU RESULT`,
                ~ case_when(.x == "EV-A71 B5"     ~ "EV-A71/CV-B5",
                            .x == "EV-A71/ RhiA"  ~ "EV-A71/RhiA",
                            .x == "EV-A71/CV A24" ~ "EV-A71/CV-A24",
                            .x == "neg"           ~ "negative",
                            .x == "CV-A8/EV-A71"  ~ "EV-A71/CV-A8",
                            .default = .x))) |>
  na.exclude()
```


#### MRI data

Reading the MRI data (i.e. lesions visible on the MRI):

```{r}
MRI <- MRI_file |>
  read_csv2(col_select = c(code, Final, Acute), show_col_types = FALSE) |> 
  rename(ID = code) |> 
  mutate(across(c("Final", "Acute"), ~ .x == "Yes"))
```

`Final` and `Acute` are not identical but it is `Final` that should be considered for
the analysis:

```{r}
filter(MRI, Final != Acute)
```

#### Bayley data

Loading the data from CliRes:

```{r}
Bayley0 <- read_excel_file(Bayley_file)
```

The tabs that we are interested in are the following:

* **CS:** cognitive scale
* **RC:** receptive communication (language scale)
* **EC:** expressive communication (language scale)
* **FM:** fine motor (motor scale)
* **GM:** gross motor (motor scale)

```{r}
Bayley_tabs <- c("CS", "RC", "EC", "FM", "GM")
```

Let's generate the data frame from these tabs:

```{r}
common_variables <- c("PARNO", "DATETESTED")

Bayley <- Bayley_tabs |> 
  map(~ c(common_variables, "SCALESCORE", .x)) |> 
  setNames(Bayley_tabs) |> 
  get_vars(Bayley0) |> 
  imap(~ rename(.x, !! paste0("SCALESCORE_", .y) := SCALESCORE)) |> 
  reduce(left_join, common_variables) |> 
  mutate(across(starts_with("DATE"), as.Date)) |> 
  rename(ParNo = PARNO) |> 
  mutate(across(ParNo, ~ stringr::str_remove(.x, "08RS_")))
```


## Preparing datasets

### Children data

#### 08RS, PCR, MRI

Generating the groups from Saras' CSV file:

```{r}
groups <- saras |> 
  select(code, group) |> 
  rename(ID = code) |> 
  unique() |> 
  arrange(ID) |> 
  mutate(across(group, ~ str_replace(.x, "Control", "control")))
```

First recoding of variables:

```{r}
recoding1 <- function(x) {
  x |>
    mutate(across(Gender, ~ c("male", "female")[.x]),
           across(starts_with("Date"), as.Date),
           across(c("Refrigerator", "AirConditioner",
                    "Motorbike", "TreatSepsis"), ~ .x < 2))
}
```

Second recoding of variables:

```{r}
recoding2 <- function(x) {
  x |>
    mutate(across(HFMD,    ~ CRF_dict$disc$GradeHFMD[.x]),
           across(MomEdu,  ~ CRF_dict$devsocsed$MomEdu[.x]),
           across(Toilet,  ~ CRF_dict$devsocsed$Toilet[.x]),
           across(Water,   ~ CRF_dict$devsocsed$Water[.x]),
           across(Outcome, ~ CRF_dict$disc$Outcome[.x]))
}
```

In the following code we are selecting and recoding the variables from the 08RS CRF,
and assigning to case or control. Note that because of the way HFMD are coded and
because we want the maximum grade, we need to twickle around a little bit:

```{r}
# because we recode grade:
CRF_dict0 <- CRF_dict
grade <- CRF_dict$disc$GradeHFMD
CRF_dict$disc$GradeHFMD <- c(last(grade), head(grade, -1))

children <- selection08RS |> 
  remove_slots("ABC") |> 
  get_vars(CRF08RS) |> 
  reduce(left_join, by = "ParNo") |> 
  rowwise() |> 
  mutate(across(contains("HFMD"), ~ c(1:6, 0)[.x] + 1),                # recode grade
         HFMD = max2(across(c(HFMDToday, HFMDAdmitted, GradeHFMD)))) |>
  ungroup() |> 
  recoding1() |> 
  recoding2() |>
  mutate(ID = as.numeric(str_remove(ParNo, "^.*-"))) |> 
  left_join(groups, "ID") |> 
  left_join(MRI, "ID") |> 
  left_join(PCR, "ID") |> 
  rename(PCR = `OUCRU RESULT`) |> 
  select(-HFMDToday, -HFMDAdmitted, -GradeHFMD, -ID) |> 
  select(ParNo, Gender, DateBirth, DateIllness, DateAdmHosp, DateAdmHTD, DateEnrol,
         DateDisc, everything()) |> 
  arrange(ParNo)

# undoing the change on the dictionary done at the top of the chunk because of recoding
CRF_dict <- CRF_dict0
rm(CRF_dict0)
```


#### Patching 02EI CRF

Conversion of IDs between 02EI and 08RS:

```{r}
(ID_conv <- tibble(s02EI = paste0("03-0", c(paste0("0", c(1, 3:9)), c("11", "13"))),
                   s08RS = paste0("03-0", c(43, 52:56, 60, 62, 78, 79))))
```

Patching the data values from the 02EI CRF:

```{r}
children <- selection02EI |> 
  get_vars(CRF02EI) |> 
  reduce(left_join, by = "studyCode") |> 
  mutate(across(IllnessDate, as.Date),
         across(c("seizures", "tracheostomy", "muscleStength", "limbParalysing",
                  "nerveParalysing"), ~ .x < 2)) |> 
  rename(ParNo         = studyCode,
         weigh         = weight,
         DateIllness   = IllnessDate,
         Seizure       = seizures,
         Trache        = tracheostomy,
         Hypertonicity = muscleStength,
         LimbPara      = limbParalysing,
         CNP           = nerveParalysing) |> 
  filter(ParNo %in% ID_conv$s02EI) |> 
  mutate(across(ParNo, ~ unname(with(ID_conv, setNames(s08RS, s02EI))[.x]))) |> 
  patch(children, "ParNo")
```


#### Stunt and waste

A function that computes stunt and waste variable from weight and height variables:

```{r}
stunt_waste <- function(x, weight, height) {
  x |> 
    mutate(age = DateEnrol - DateBirth,
           z   = anthro::anthro_zscores(c(male = 1, female = 2)[Gender],
                                        as.numeric(age),
                                        weight = {{ weight }},
                                        lenhei = {{ height }})[c("zlen", "zwfl")]) |> 
    unnest(z) |> 
    mutate(stunting = zlen < -2,
           wasting  = case_when(is.na(zwfl) ~ NA_character_,
                                zwfl < -3 ~ "severe",
                                zwfl < -2 ~ "moderate",
                                .default = "no")) |> 
    select(-c(zlen, zwfl))
}
```

Adding stunt and waste data to the `children` data frame:

```{r}
children <- stunt_waste(children, weigh, height)
```


#### Missing values

A function that computes the numbers and percentages of missing values per variable:

```{r}
number_of_NA <- function(x) {
  x |> 
    select(- group) |> 
    map_dfr(~ sum(is.na(.x))) |> 
    pivot_longer(! ParNo, names_to = "variable", values_to = "number_of_NA") |> 
    mutate(percentage_of_NA = round(100 * number_of_NA / nrow(x))) |> 
    select(- ParNo)
}
```

Let's look at the missing values among cases:

```{r}
children |>
  filter(group == "HFMD") |> 
  number_of_NA() |> 
  print(n = Inf)
```

And among controls:

```{r}
children |>
  filter(, group == "control") |> 
  number_of_NA() |> 
  filter(percentage_of_NA < 100) |> 
  print(n = Inf)
```

A function that displays the `NA` values:

```{r}
show_NA <- function(x) {
  cases <- filter(children, group == "HFMD")
  controls <- filter(children, group == "control")
  left_join(number_of_NA(cases),
            number_of_NA(controls) |>
              filter(percentage_of_NA <100),
            "variable") |> 
    print(n = Inf)
}
```

Putting the cases and controls together:

```{r}
show_NA(children)
```

Retrieving the height, weight and stunt data from Saras' CSV file:

```{r}
saras_stunt <- saras |> 
  select(code, WEIGHT, HEIGHT, headCircumference, stunt, waste) |> 
  unique() |> 
  rename(ParNo = code,
         headCircumference_S = headCircumference) |> 
  mutate(across(ParNo, ~ paste0(ifelse(.x < 923, "03-", "05-"), sprintf("%03d", .x))))
```

Merging with our children data frame for comparison:

```{r}
comparing_data <- children |> 
  select(ParNo, headCircumference, height, weigh, stunting, wasting) |> 
  left_join(saras_stunt, "ParNo")
```

Checking that children codes match:

```{r}
setdiff(saras_stunt$ParNo, comparing_data$ParNo)
setdiff(comparing_data$ParNo, saras_stunt$ParNo)
```

A function that plots the values from the two data sources:

```{r}
plot_comp <- function(...) {
  plot(..., col = adjustcolor(4, .15), pch = 19,
       xlab = "CRF", ylab = "Saras'", asp = 1)
  abline(0, 1, col = 2)
}
```

Comparing height and weight data:

```{r fig_comparisons, fig.width = 6, fig.height = 3}
opar <- par(mfrow = 1:2, pty = "s", plt = rep(c(.2, .93), 2), bty = "o")

comparing_data |> 
  filter_out(is.na(height)) |> 
  with(plot_comp(height, HEIGHT))
mtext("heights (cm)", line = .2)

comparing_data |> 
  filter_out(is.na(weigh)) |> 
  with(plot_comp(weigh, WEIGHT))
mtext("weights (kg)", line = .2)

par(opar)
```

Checking whether there are any missing values in Saras' file that are not missing in
the CRF:

```{r}
filter(comparing_data, is.na(headCircumference_S), ! is.na(headCircumference))
filter(comparing_data, is.na(HEIGHT),              ! is.na(height))
filter(comparing_data, is.na(WEIGHT),              ! is.na(weigh))
filter(comparing_data, is.na(stunt),               ! is.na(stunting))
filter(comparing_data, is.na(waste),               ! is.na(wasting))
```

Checking whether some of the stunt and waste variables in Saras' file couldn't be
recomputed:

```{r}
filter(saras_stunt, (is.na(WEIGHT) | is.na(HEIGHT)) & ! is.na(stunt))
filter(saras_stunt, (is.na(WEIGHT) | is.na(HEIGHT)) & ! is.na(waste))
```

Comparing stunt data:

```{r}
comparing_data |> 
  mutate(across(stunt, ~ .x == "Stunt")) |> 
  filter_out(is.na(stunting)) |> 
  group_by(stunting, stunt) |> 
  tally() |> 
  ungroup() |> 
  setNames(c("CRF", "Saras'", "n"))
```

Patching the `children` data frame with Saras' CSV file for weight, height, stunt and
waste, and recomputing our stunting and wasting data:

```{r}
children <- children |> 
  left_join(select(saras_stunt, -c(stunt, waste)), "ParNo") |> 
  mutate(weigh             = WEIGHT,
         height            = HEIGHT,
         headCircumference = headCircumference_S) |> 
  select(-c(WEIGHT, HEIGHT, headCircumference_S)) |> 
  stunt_waste(weigh, height)
```

Here is the new situation regarding missing values:

```{r}
show_NA(children)
```


### Follow-up data

A function that generates the time points:

```{r}
make_time_points <- function(x) {
  children |> 
    select(ParNo, DateEnrol, DateDisc) |> 
    left_join(x, "ParNo") |> 
    mutate(time_diff = DateTested - DateEnrol,
           time1 = 0, time2 = 6, time3 = 18, # in months
           across(c(time1, time2, time3), ~ as.numeric(abs(time_diff - 30 * .x)))) |> 
    rowwise() |> 
    mutate(min_delay = min(across(c(time1, time2, time3)))) |> 
    ungroup() |> 
    mutate(time_point = ifelse(min_delay == time1,
                               "enrollment", ifelse(min_delay == time2,
                                                    "6 months", "18 months")))
}
```

A function that gets the IDs of children with duplicated assessments:

```{r}
get_IDs_with_duplicated <- function(x) {
  x |> 
    filter(! is.na(time_point)) |> 
    group_by(ParNo) |> 
    group_modify(~ .x |>
                   group_by(time_point) |>
                   tally()) |> 
    ungroup() |> 
    filter(n > 1) |> 
    pull(ParNo) |> 
    unique()
}
```

A function that uses the previous two to generate the data with duplicated assessments:

```{r}
show_duplicated_assessments <- function(x) {
  data_with_time_points <- make_time_points(x)
  IDs_with_duplicates <- get_IDs_with_duplicated(data_with_time_points)
  filter(data_with_time_points, ParNo %in% IDs_with_duplicates)
}
```


#### M-ABC data

```{r}
ABC <- CRF08RS$ABC |> 
  select(ParNo, DateTested, ends_with("ISS")) |> 
  mutate(across(starts_with("Date"), as.Date)) |> 
  arrange(ParNo, DateTested)
```

Of note, here

* `MD` stands for manual dexterity,
* `AC` stands for aiming and catching and
* `BAL` stands for balance.

```{r}
ABC |> 
  na.exclude()
```

Write duplicates to disk:

```{r}
ABC |>
  show_duplicated_assessments() |> 
  writexl::write_xlsx("M-ABC2.xlsx")
```

Here all the duplicates are complete. We'll simply keep all the earlier ones:

```{r}
ABC2 <- ABC |> 
  make_time_points() |>
  arrange(ParNo, time_point, min_delay) |> 
  group_by(ParNo, time_point) |> 
  group_modify(~ head(.x, 1)) |> 
  ungroup() |> 
  select(-DateEnrol, -DateDisc, -min_delay, - time_diff, -time1, -time2, -time3) |> 
  rename(Date_ABC = DateTested)
```


#### Bayley data

```{r}
Bayley <- rename(Bayley, DateTested = DATETESTED)

Bayley |>
  show_duplicated_assessments() |> 
  writexl::write_xlsx("Bayley2.xlsx")
```

This shows that

* there is one and only one complete measurement per time point
* the complete measurement is always the earlier one, except for patient `03-514`

In consequence, we decide to simply filter out all the incomplete duplicates:

```{r}
Bayley2 <- Bayley |> 
  make_time_points() |>
  group_by(ParNo, time_point) |> 
  group_modify(~ {if (nrow(.x) > 1) return(na.exclude(.x)); .x }) |> 
  ungroup() |> 
  select(-DateEnrol, -DateDisc, -min_delay, - time_diff, -time1, -time2, -time3) |> 
  rename(Date_Bayley = DateTested)
```


#### Merging

Merging the M-ABC and Bayley data:

```{r}
followups <- full_join(ABC2, Bayley2, c("ParNo", "time_point"))
```


#### Visualization

A function that prepends all the data frames of a list `x` of data frames with `n`
columns of the `v` values:

```{r}
prepend_white <- function(x, n, v) {
  nbrows <- nrow(x[[1]])
  white_space <- v |> 
    rep(n * nbrows) |> 
    matrix(nbrows) |> 
    as.data.frame()
  map(x, ~ cbind(white_space, .x))
}
```

A function that (i) splits the data frame `x` into a list of data frame of `n` rows 
(except possibly for the last slot), (ii) prepends each of these data frames with `wc`
columns of 1s, and (iii) concatenate all these data frames side by side into a matrix:

```{r}
side_by_side <- function(x, n, wc) {
  x |>
    select(-ParNo) |>
    split_df(n) |> 
    append_last() |>
    prepend_white(wc, 1) |> 
    reduce(cbind) |> 
    as.matrix()
}
```

A tuning of `image2()`:

```{r}
image3 <- function(x, col_no, col_yes) {
  image2(0:ncol(x), 0:nrow(x), x, axes = FALSE, ann = FALSE, col = c(col_no, col_yes))
}
```

The function that plots the heatmap:

```{r}
heatmap2 <- function(x, nbrow = 45, nb_wc = 2,
                     col_Bayley = adjustcolor("red", .3),
                     col_ABC    = adjustcolor("blue", .3),
                     col_NA     = adjustcolor("white", 0),
                     col_lines  = "white", cx = .5) {
# plotting M-ABC data:
  tmp <- x |> 
    select(-Date_Bayley) |> 
    pivot_wider(names_from = time_point, values_from = Date_ABC) |> 
    side_by_side(nbrow, nb_wc)
  image3(tmp, col_NA, col_Bayley)
  
# adding Bayley data:
  par(new = TRUE)
  x |> 
    select(-Date_ABC) |> 
    pivot_wider(names_from = time_point, values_from = Date_Bayley) |> 
    side_by_side(nbrow, nb_wc) |> 
    image3(col_NA, col_ABC)

# adding separation lines:
  abline(v = 0:ncol(tmp), col = col_lines)
  abline(h = 0:nbrow, col = col_lines)

# adding children IDs:
  ids <- str_remove(unique(x$ParNo), "^.*-")
  sel <- 1:length(ids)
  by <- 3 + nb_wc
  ncol_tmp <- ncol(tmp)
  nbcol <- ncol_tmp / by
  xs <- rep(seq(1, ncol_tmp, by), each = nbrow)[sel]
  ys <- rep(rev(1:nbrow - .5), nbcol)[sel]
  text(xs, ys, ids, cex = cx)

# adding time points:
  xx <- seq(1 + nb_wc, ncol_tmp, by) - .5
  mtext(rep(c("e", "1", "2"), nbcol)[sel], at = sort(c(xx, xx + 1, xx + 2)), cex = cx)
}
```

An overview of the M-ABC and Bayley data for all the children and the 3 time points:

```{r fig_followup_data_availability, fig.width  = 7, fig.height = 5.5}
opar <- par(plt = c(.07, .97, .07, .95))

expand_grid(ParNo      = unique(followups$ParNo),
            time_point = c("enrollment", "6 months", "18 months")) |> 
  left_join(followups, c("ParNo", "time_point")) |> 
  select(ParNo, time_point, starts_with("Date")) |> 
  mutate(across(starts_with("Date"), ~ as.numeric(! is.na(.x)) + 1)) |> 
  heatmap2()

color_codes <- rgb(c(255, 191, 192),
                   c(193, 194, 147),
                   c(193, 255, 213), names = c("m-abc", "bayley", "both"), max = 255)

usr <- par("usr")
legend(usr[1] + mean(usr[1:2]), usr[3] - 2, legend = c("M-ABC", "Bayley", "both"),
       fill = color_codes, bty = "n", horiz = TRUE, xpd = TRUE, xjust = .5, yjust = .5)

par(opar)
```

where blue is where the Bayley data are available, red is where the M-ABC data are
available, purple is where both data are available and white is where none of the data
are available.


## Explorations


### EV-A71 and severity

A function that computes a summary table of HFMD cases according to the variable `var`:

```{r}
make_table <- function(df, var) {
  df |> 
    filter(group == "HFMD") |> 
    group_by({{ var }}) |> 
    count() |> 
    ungroup() |> 
    mutate(`p (%)` = round(100 * n / sum(n)))
}
```

The various PCR results:

```{r}
virus_table_all <- children |>
  make_table(PCR) |> 
  rename(virus = PCR)

print(virus_table_all, n = Inf)
```

Severity:

```{r}
(severity_table_all <- children |>
  make_table(HFMD) |> 
  rename(severity = HFMD))
```

Severity vs EV71:

```{r}
children |> 
  filter(group == "HFMD") |>
  mutate(EV71 = str_detect(PCR, "EV-A71")) |> 
  with(table(HFMD, EV71))
```

A function that plots the proportion of `TRUE` in the variable `y` of the data frame
`x` as a function of severity grade:

```{r}
plot_severity <- function(df, ...) {
  df |> 
    group_by(HFMD, y) |> 
    count() |> 
    ungroup() |> 
    na.exclude() |> 
    pivot_wider(names_from = y, values_from = n) |>
    mutate(across(everything(), ~ replace_na(.x, 0))) |> 
    rename(x = `TRUE`) |>
    mutate(at         = as.numeric(as.factor(HFMD)),
           n          = x + `FALSE`,
           binom_test = map2(x, n, binom.test),
           estimate   = map_dbl(binom_test, ~ .x$estimate),
           confint    = map(binom_test, ~ setNames(.x$conf.int, c("lb", "ub")))) |> 
    unnest_wider(confint) |> 
    with({
      plot(at, estimate, ylim = 0:1, axes = FALSE, xlab = "severity grade",
           type = "n", ...)
      
      v <- na.exclude(df$y)
      btest <- binom.test(sum(v), length(v))
      confint <- btest$conf.int
      rect(min(at) - 1, confint[1], max(at) + 1, confint[2], col = "lightgrey",
           border = NA)
      abline(h = btest$estimate, lty = 2, lwd = 2, col = "grey")

      points(at, estimate, pch = 19, col = 4)
      arrows(at, lb, at, ub, .1, 90, 3, col = 4, lwd = 2)
      axis(1, at, str_remove(HFMD, "grade "))
      axis(2)
    })
}
```

```{r}
virusXseverity <- function(virus) {
  children |> 
    filter(group == "HFMD") |> # (not necessary)
    mutate(y = str_detect(PCR, virus)) |> 
    plot_severity(ylab = paste("proportion of", virus))
}
```

The proportion of EV71 as a function of the severity grade:

```{r fig_EV71_severity}
virusXseverity("EV-A71")
```

```{r fig_EV_severity}
virusXseverity("EV")
```

```{r fig_EV$_severity}
virusXseverity("EV$")
```

```{r fig_CVA6_severity}
virusXseverity("CV-A6")
```

```{r fig_CVA10_severity}
virusXseverity("CV-A10")
```

The proportion of the presence of lesions observed by MRI as a function of the severity
grade:

```{r fig_lesions_severity}
children |> 
  rename(y = Final) |> 
  plot_severity(ylab = "proportion of MRI with lesions")
```

A function that filters the patients in the `followups` data frame according to some
rules on the variables on the `children` data frame:

```{r}
filter_patients <- function(...) {
  sel <- children |> 
    filter(...) |> 
    pull(ParNo)
  filter(followups, ParNo %in% sel)
}
```


### MRI

```{r}
virus_table_mri <- children |> 
  filter_out(is.na(Final)) |>
  make_table(PCR) |> 
  rename(virus = PCR)
```

```{r}
virus_table <- left_join(virus_table_all, virus_table_mri, by = "virus") |> 
  mutate(across(where(is.numeric), ~replace_na(.x, 0)))

print(virus_table, n = Inf)
```

```{r fig_MRI_nonMRI}
with(virus_table, plot(n.x, n.y))
```

```{r}
severity_table_mri <- children |>
  filter_out(is.na(Final)) |>
  make_table(HFMD) |> 
  rename(severity = HFMD)
```

```{r}
(severity_table <- left_join(severity_table_all, severity_table_mri,
                             by = "severity") |> 
  mutate(across(where(is.numeric), ~replace_na(.x, 0))))
```

```{r}
tmp <- children |>
  filter(group == "HFMD") |> 
  select(Final, PCR, HFMD) |> 
  mutate(across(Final, ~ ! is.na(.x)),
         across(HFMD, ~ ifelse(is.na(.x) | .x == "Not Applicable", "grade 1", .))) |> 
  na.exclude()
```

```{r fig_probaMRI_severity}
children |>
  filter(group == "HFMD") |> 
  select(Final, HFMD) |> 
  mutate(y = ! is.na(Final)) |> 
  plot_severity(ylab = "proportion of MRI done")
```

```{r}
children |>
  filter(group == "HFMD") |> 
  select(Final, PCR, HFMD) |> 
  mutate(across(Final, ~ ! is.na(.x)),
         across(PCR, ~ str_detect(.x, "EV-A71")),
         across(HFMD, ~ ifelse(is.na(.x) | .x == "Not Applicable", "grade 1", .x))) |> 
  na.exclude()
```

```{r}
children |> 
  select(Final, Acute) |> 
  na.exclude() |> 
  group_by(Final, Acute) |> 
  count()
```

```{r}
children |> 
  group_by(is.na(Final), is.na(Acute)) |> 
  count()
```

```{r}
children |> 
  filter_out(is.na(Final)) |> 
  select(Final, PCR) |> 
  mutate(EV71 = str_detect(PCR, "EV-A71")) |> 
  with(table(Final, EV71))
```


### Scores

```{r}
hist3 <- function(x, ...) {
  out <- hist(x, (min(x, na.rm = TRUE) - 1):(max(x, na.rm = TRUE) + 1), ...)
  out$breaks <- out$breaks + .5
  out$mids <- out$mids + .5
  out
}
```

```{r}
plot_score_distr <- function(df) {
  opar <- par(mfrow = c(5, 3), cex = 1, plt = c(.2, .95, .3, .94))
  df |>
    select(ends_with("ISS"), "CS", "RC", "EC", "FM", "GM") |> 
    iwalk(~ {
      hdata <- hist3(.x, plot = FALSE)
      ddata <- density(na.exclude(.x))
      plot(hdata, main = NA, freq = FALSE, ylim = c(0, max(hdata$density, ddata$y)),
           xlab = paste(.y, "score"), ylab = "density")
      lines(ddata, col = 2, lwd = 2)
    })
  par(opar)
}
```

```{r fig_score_distributions, fig.height = 5 * figh / 3, fig.width = figw}
plot_score_distr(followups)
```

```{r fig_score_distr_timepoints, fig.height = 5 * figh / 3, fig.width = figw}
c("enrollment", "6 months", "18 months") |> 
  walk(~ followups |> 
         filter(time_point == .x) |> 
         plot_score_distr())
```

```{r}
plot_score_distr2 <- function(df) {
  opar <- par(mfrow = c(5, 3), cex = 1, plt = c(.2, .95, .3, .94))
  df |>
    select(ends_with("ISS"), starts_with("SCALE")) |> 
    iwalk(~ {
      hdata <- hist3(.x, plot = FALSE)
      ddata <- density(na.exclude(.x))
      plot(hdata, main = NA, freq = FALSE, ylim = c(0, max(hdata$density, ddata$y)),
           xlab = paste(.y, "score"), ylab = "density")
      lines(ddata, col = 2, lwd = 2)
    })
  par(opar)
}
```

```{r fig_score_distributions2, fig.height = 5 * figh / 3, fig.width = figw}
plot_score_distr2(followups)
```

```{r fig_score_distr_timepoints2, fig.height = 5 * figh / 3, fig.width = figw}
c("enrollment", "6 months", "18 months") |> 
  walk(~ followups |> 
         filter(time_point == .x) |> 
         plot_score_distr2())
```

```{r}
mean_sd <- function(df) {
  df |>
    map(~ tibble(mean = mean(.x, na.rm = TRUE), sd = sd(.x, na.rm = TRUE))) |> 
    list_rbind(names_to = "score")
}
```

```{r}
followups |> 
  select(CS, RC, EC, FM, GM) |> 
  mean_sd()
```

```{r}
followups |> 
  select(starts_with("SCALE")) |> 
  mean_sd()
```


#### RC

```{r}
fu2 <- followups |> 
  mutate(date = ifelse(is.na(Date_ABC), Date_Bayley, Date_ABC)) |> 
  select(-starts_with("Date_")) |> 
  pivot_longer(-c(ParNo, time_point, date)) |> 
  na.exclude() |> 
  left_join(children, "ParNo") |> 
  select(ParNo, time_point, date, name, value, Gender, DateBirth, DateIllness, MomEdu,
         HFMD, group, Final, PCR, age, stunting, wasting)
```

```{r}
tps <- c("enrollment", "6 months", "18 months")
fu2 |> 
  filter(name == "RC") |>
  select(time_point, value) |> 
  mutate(across(time_point, ~ setNames(1:3, tps)[time_point])) |> 
  group_by(time_point) |> 
  group_walk(~ hist3(.x$value, xlim = c(2, 45)))
```

```{r}
fu3 <- fu2 |> 
  mutate(EV71 = str_detect(PCR, "EV-A71"),
         across(c(EV71, HFMD),
                ~ ifelse(group == "control", "not done", as.character(.x)))) |> 
  select(-c(ParNo, date, DateBirth, DateIllness, Final, PCR)) |> 
  na.exclude()
  
#fu3 |> 
#  filter(name == "RC") |> 
#  select(-name) |> 
#  randomForest(value ~ ., data = _) |> 
#  importance() |> 
#  as.data.frame() |> 
#  rownames_to_column("variable") |> 
#  arrange(desc(IncNodePurity))
```

```{r}
variable_importance <- function(var) {
  fu3 |> 
    filter(name == var) |> 
    select(-name) |> 
    randomForest(value ~ ., data = _) |> 
    importance() |> 
    as.data.frame() |> 
    rownames_to_column("variable") |> 
    arrange(desc(IncNodePurity))
} 
```

```{r}
followups |>
  select(ends_with("ISS"), "CS", "RC", "EC", "FM", "GM") |> 
  names() |> 
  set_names() |> 
  map(variable_importance)
```

```{r}
fu2 |> 
  group_by(name, time_point, group) |> 
  tally() |> 
  ungroup() |> 
  pivot_wider(names_from = group, values_from = n) |> 
  print(n = Inf)
```

```{r}
fu2 |> 
  filter(name == "BAL3ISS") |> 
  with(plot(age, value))
```

```{r}
a <- followups |> 
  select(SCALESCORE_CS:GM) |> 
  pivot_longer(everything()) |> 
  mutate(variable = ifelse(str_starts(name, "SCALE"), "scale", "score"),
         across(name, ~ paste(str_remove(.x, "SCALESCORE_"))),
         group = rep(1:(n() / 2), each = 2),
         name = paste(name, group)) |> 
  pivot_wider(id_cols = -group, names_from = variable, values_from = value) |> 
  na.exclude()

with(a, plot(jitter(score), jitter(scale), xlab = "raw score", ylab = "scaled score"))
with(a, plot(jitter(score), jitter(scale), xlab = "raw score", ylab = "scaled score",
             ylim = c(0, 20)))


a |> 
  filter(str_starts(name, "CS")) |> 
  with(plot(jitter(score), jitter(scale), xlab = "raw score", ylab = "scaled score"))
```

```{r}
c("CS", "RC", "EC", "FM", "GM") |> 
  walk(~ a |> 
         filter(str_starts(name, .x)) |> 
         with(plot(jitter(score), jitter(scale),
                   xlab = "raw score", ylab = "scaled score")))
```

```{r}
a |> 
  filter(str_starts(name, "CS")) |> 
  with({
    hist(scale)
    hist(score)
  })
```




### HFMD vs controls

From here we work with 2 data frames: `children` that contains the children
information, and `followups` that contains the follow-up data. Number of missing data
points:

```{r}
children |> 
  select(group, age, MomEdu, Gender, stunting) |> 
  map_int(~ sum(is.na(.x)))
```

Some common code:

```{r}
cols <- c(2, 4)
adjcol <- function(...) adjustcolor(..., alpha = .3)

barplot2 <- function(height, x) {
  barplot(height, names.arg = x, beside = TRUE, ylab = "proportion",
          col = adjcol(rev(cols)))
}

add_legend1 <- function(where = "topright") {
  legend(where, legend = c("control", "HFMD"), fill = adjcol(rev(cols)), bty = "n")
}
```

Let's look at the age distribution of the controls and HFMD cases:

```{r fig_age_distribution, fig.width = eps * figw, fig.height = eps * figh}
tmp <- children |>
  mutate(across(age, ~ as.numeric(.x) / 30)) |> 
  group_by(group) |> 
  group_map(~ .x |> pull(age) |> density(from = 0))

tmp |> 
  map(~ .x |> unclass() |> magrittr::extract(c("x", "y")) |> as_tibble()) |> 
  bind_rows() |> 
  map(range) |> 
  with(plot(NA, xlim = x, ylim = y, xlab = "age (months)", ylab = "density"))

tmp |> 
  walk2(cols, ~ polygon(adding_zero_ys(.x), col = adjcol(.y), border = NA))

add_legend1()
```

Let's now look at the level of the mother's education:

```{r fig_mom_edu, fig.width = eps * figw, fig.height = eps * figh}
edu_order <- c("Never been to school", "Attended some primary school",
               "Completed primary school (5th gr)",
               "Completed lower secondary school (9th gr)",
               "Completed higher secondary school (12th gr)",
               "Completed university/college degree", "Completed postgraduate degree")

children |> 
  with(table(MomEdu, group)) |>
  prop.table(2) %>%
  `[`(edu_order, ) |> 
  t() |> 
  barplot2(c("none", "some", "5th g", "9th g", "12th g", "collg", "postg"))

add_legend1()
```

Stunt and gender:

```{r fig_sex_stunt, fig.width = eps * figw, fig.height = eps * figh}
children |> 
  select(group, Gender, stunting) |> 
  na.exclude() |> 
  group_by(group) |> 
  group_split() |> 
  map(~ .x |> group_by(Gender, stunting) |> tally() |> ungroup()) |> 
  reduce(left_join, c("Gender", "stunting")) |> 
  select(starts_with("n")) |> 
  as.matrix() |> 
  prop.table(2) |> 
  t() %>%
  `[`(2:1, ) |> 
  barplot2(rep(c("female", "male"), each = 2))

mtext(rep(c("non-stunt", "stunt"), 2), 1, 1.5, at = seq(2, 11, 3))
add_legend1()
```


### Age distributions

```{r}
ages <- followups |> 
  left_join(select(children, ParNo, DateBirth), "ParNo") |> 
  mutate(age_ABC = time_length(Date_ABC - DateBirth, "year"),
         age_Bayley = time_length(Date_Bayley - DateBirth, "year"),
         age = ifelse(is.na(age_ABC), age_Bayley, age_ABC)) |> 
  select(ParNo, time_point, age) |> 
  na.exclude()
```

```{r fig_age_distributions, fig.height = 3 * eps * figh}
opar <- par(mfrow = c(3, 1), cex = 1)
walk(c("enrollment", "6 months", "18 months"), ~ {
       filter(ages, time_point == .x) |> 
       pull(age) |> 
       hist(xlim = c(0, 6), ylim = c(0, 135), main = NA, yaxs = "i",
            xlab = "age (years)", ylab = "counts")
       text(3, 130, .x)
  })
par(opar)
```


### Timings of followups

```{r}
btw_fup <- ages |> 
  pivot_wider(names_from = time_point, values_from = age) |> 
  mutate(fu1 = 12 * (`6 months` - enrollment),
         fu2 = 12 * (`18 months` - enrollment),
         fu3 = 12 * (`18 months` - `6 months`))
```

```{r timings_followups}
fuh <- btw_fup |> 
  select(starts_with("fu")) |> 
  map(hist, plot = FALSE)

plot(fuh[[1]], xlab = "timing of follow-up (months)", ylab = "counts", main = NA,
     xlim = c(0, max(unlist(map(fuh, "breaks")))), yaxs = "i", col = 4,
     ylim = c(0, 1.04 * max(unlist(map(fuh, "counts")))))
plot(fuh[[2]], col = 4, add = TRUE)
abline(v = c(6, 18), col = 2, lwd = 3)

plot(fuh[[3]], add = TRUE)
abline(v = 12, lwd = 3)
```




## Analysis per time point

### Propensity scores

All the covariates when considering the controls as the controls:

```{r}
covariables_all <- select(children, ParNo, Gender, age, MomEdu, stunting, group)
```

The covariates for the HFMD cases only and when considering the grades 2a as the
controls:

```{r}
covariables_sev <- children |>
  filter(group == "HFMD") |> 
  mutate(group = ifelse(HFMD == "grade 2a", "control", "HFMD")) |> 
  select(ParNo, Gender, age, MomEdu, stunting, group)
```

The covariates for the HFMD cases only and when considering the non EV-A71 as the
controls:

```{r}
covariables_EV71 <- children |>
  filter(group == "HFMD") |> 
  mutate(group = ifelse(str_detect(PCR, "EV-A71"), "HFMD", "control")) |> 
  select(ParNo, Gender, age, MomEdu, stunting, group)
```

The covariates for the HFMD cases with an MRI only and when considering those without
any lesions as the controls:

```{r}
covariables_MRI <- children |>
#  filter(group == "HFMD") |> 
  filter_out(is.na(Final)) |> 
  mutate(group = ifelse(Final, "HFMD", "control")) |> 
  select(ParNo, Gender, age, MomEdu, stunting, group)
```

The formula:

```{r}
ps_formula <- group ~ Gender + age + MomEdu + stunting
```

A function that generates a data set for a given response variable at a given time
point:

```{r}
make_data <- function(data, response, timepoint, covariab = covariables) {
  data |> 
    filter(time_point == timepoint) |> 
    select(ParNo, all_of(response)) |> 
    na.exclude() |> 
    left_join(covariab, "ParNo") |> # here we can specify the covariables we want
    select(-ParNo) |> 
    select(all_of(response), group, everything()) |> 
    na.exclude()
}
```

The scores:

```{r}
scores <- c("MD1ISS", "MD2ISS", "MD3ISS", "AC1ISS", "AC2ISS", "BAL1ISS", "BAL2ISS",
            "BAL3ISS", "CS", "RC", "EC", "FM", "GM")
```

A function that prepares the follow-up data:

```{r}
prepare_fu_data <- function(data = followups, min_nb = 4, covariab = covariables) {
  expand_grid(response  = scores,
              timepoint = c("enrollment", "6 months", "18 months")) |> 
    mutate(dataset = map2(response, timepoint, ~ make_data2(data, .x, .y, covariab)),
           ctable  = map(dataset, ~ as.vector(table(.x$group)))) |> 
    filter(map_lgl(ctable, ~ length(.x) > 1)) |> 
    mutate(ctable = map(ctable, ~ setNames(.x, c("control", "HFMD")))) |> 
    unnest_wider(ctable) |> 
    filter(HFMD > min_nb)
}
```

A function that generates ATT (Average Treatment effect on Treated) weights from a
logistic regression:

```{r}
logistic_ps <- function(data) {
  data |>
    mutate(across(group, ~ .x == "HFMD"),
           weights = ps_formula |>
             glm(binomial, data = pick(everything())) |> 
             predict(type = "response") %>%
             { ifelse(group, 1, . / (1 - .)) }) |>  # ATT 
    pull(weights)
}
```

A function that generates weights from the `twang` package:

```{r}
twang_ps <- function(data, ...) {
  data |> 
    mutate(across(group, ~ as.integer(group == "HFMD")),
           across(where(~ is.character(.x) | is.logical(.x)), as.factor),
           across(age, as.numeric)) |> 
    as.data.frame() |> 
    ps(ps_formula, data = _, estimand = "ATT", stop.method = "es.mean",
       verbose = FALSE, ...) |> 
    get.weights("es.mean")  
}
```

A function that computes the TMLE:

```{r}
default_libs <- c("SL.glm", "SL.glmnet", "SL.mean", "SL.gam")

tmle2 <- function(data, response, family = "gaussian",
                  Qlib = default_libs, glib = default_libs, ...) {
  data |> 
    select(Gender, age, MomEdu, stunting) |> 
    mutate(across(where(~ is.character(.x) | is.logical(.x)), as.factor),
           across(age, as.numeric)) |>
    tmle(data[[response]], as.integer(data$group == "HFMD"), W = _, family = family,
         Q.SL.library = Qlib, g.SL.library = glib) |> 
    mextract2(c("estimates", "ATT")) |> 
    mextract(c("psi", "CI", "pvalue")) |> 
    unlist() |> 
    setNames(c("cf_tm", "2.5 %tm", "97.5 %tm", "p_tm"))
}
```

A function that generates the formula of the model with the response `y`:

```{r}
reformulate2 <- function(y) {
  reformulate(c("Gender", "age", "MomEdu", "stunting", "group"), y)
}
```

A function that processes the output of a model in the `pipeline()` function that
follows:

```{r}
process_model <- function(x, model, nm) {
  x |> 
    mutate(!!paste0("cf_", nm) := map_dbl({{ model }}, coef2),
           ci                   = map({{ model }}, confint2)) |>   
    unnest_wider(ci) |> 
    rename_with(~ paste0(.x, nm), ends_with("%")) |> 
    mutate(!!paste0("p_", nm) := map_dbl({{ model}}, get_p))
}
```

The pipeline:

```{r}
pipeline <- function(data = followups, min_nb = 4, covariab = covariables) {
  data |> 
    prepare_fu_data2(min_nb, covariab) |> 
    mutate(dataset = map(dataset,
                         ~ mutate(.x, ow = logistic_ps(pick(everything())),
                                      tw = twang_ps(pick(everything())),
                                      t2 = twang_ps(pick(everything()),
                                                    n.trees = 5000,
                                                    interaction.depth = 2))),
           formula = map(response, ~ reformulate("group", .x)),
           glm0 = map2(formula,  dataset, ~ glm(.x, data = .y)),
           glm1 = map2(response, dataset, ~ glm(reformulate2(.x), data = .y)),
           m_ow = map2(formula,  dataset, ~ svyglm2(.x, .y, ~ow)),
           m_tw = map2(formula,  dataset, ~ svyglm2(.x, .y, ~tw)),
           m_t2 = map2(formula,  dataset, ~ svyglm2(.x, .y, ~t2))) |> 
    process_model(glm0, "l0") |> 
    process_model(glm1, "l1") |> 
    process_model(m_ow, "ow") |> 
    process_model(m_tw, "tw") |> 
    process_model(m_t2, "t2") #|> 
# 6 - TMLE:           
#    mutate(tmle = map2_dfr(dataset, response, tmle2)) |> 
#    unnest_wider(tmle)
}
```

### Diagnostics

A function that computes balance statistics for diagnostic of the propensity scores:

```{r}
bal_tab <- function(data, w) {
  data |> 
    mutate(across(group, ~ .x == "HFMD")) |> 
    bal.tab(ps_formula, data = _, weights = w, s.d.denom = "treated")
}
```

A function that retrieves the balance adjustments for diagnostic of the propensity
scores:

```{r}
get_diff <- function(x) x$Balance$Diff.Adj
```

A function that produces the maximum values of the balance adjustments for diagnostic
of the propensity scores:

```{r}
get_max_diff <- function(data, w) {
  data |> 
    bal_tab(w) |> 
    get_diff() |> 
    max()
}
```

A function that computes the diagnostic metrics for a dataset `x` and for the
propensity score weight `psw`:

```{r}
diagnostic_metrics <- function(x, psw) {
  y <- x[[psw]]
  y <- c(max = max,
         mea = mean,
         med = median,
         ess = function(x) sum(x)^2 / sum(x^2)) |>
    map_dfc(~ .x(y)) |> 
    mutate(dif = abs(mea - med),
           bal = get_max_diff(x, psw))
  names(y) <- paste0(psw, "_", names(y))
  y
}
```

A function that runs `diagnostic_metrics()` on all the datasets of an output `x` of the
`pipeline()` function, and for the propensity score weight `psw`:

```{r}
diagnostic_ps <- function(x, psw) {
  x |> 
    select(response, timepoint, control, HFMD, dataset) |> 
    mutate(diag = map(dataset, diagnostic_metrics, psw)) |> 
    unnest_wider(diag) |> 
    mutate(n = control + HFMD) |> 
    select(-c(dataset, control, HFMD))
}
```


### Plotting effects

The methods and their color code:

```{r}
methods <- c("l0", "l1", "ow", "tw", "t2")
cols <- RColorBrewer::brewer.pal(length(methods), "Set1")
```

The time points:

```{r}
timepoints <- c("enrollment", "6 months", "18 months")
template <- tibble(timepoint = timepoints)
```

A function that prepares the data:

```{r}
prepare_data <- function(method, data) {
  data |> 
    select(timepoint, ends_with(method), -starts_with(c("p", "m"))) |> 
    left_join(template, y = _, "timepoint") |> 
    mutate(x = 1:n()) |> 
    setNames(c("timepoint", "y", "l", "u", "x"))
}
```

A function that plots the effect for 1 score, for the 3 time points and the 5 methods:

```{r}
plot_effects <- function(x) {
  x |>
    map2(seq(-.2, .2, .1), ~ mutate(.x, x = x + .y)) |> 
    map2(cols, ~ mutate(.x, col = .y)) |> 
    bind_rows() |> 
    with({
      plot(x, y, ylim = c(min(l, na.rm = TRUE), max(u, na.rm = TRUE)), col = col,
           axes = FALSE, xlab = NA, pch = 19, ylab = "effect of disease on score")
      abline(h = 0)
      arrows(x, l, x, u, .03, 90, 3, col)
      })
  axis(2)
}
```

A function that combines `prepare_data()` and `plot_effects()`:

```{r}
plot_1_variable <- function(var, data) {
  methods |> 
    map(prepare_data, filter(data, response == var)) |> 
    plot_effects()

  mtext2(paste(var, "score"))
  mtext2(timepoints, 1, at = seq_along(timepoints))
}
```

The function that plots the effects for all the scores, for all time points and
methods:

```{r}
plot_effects_all <- function(x) {
  plot(1, type = "n", ann = FALSE, axes = FALSE)
  legend("topleft", col = cols, lwd = 1, pch = 19, bty = "n",
         legend = c("no correction", "multivariate model", "PS by logistic regression",
                    "PS by gradient boosting, default",
                    "PS by gradient boosting, Saras'"))
  x |> 
    pull(response) |> 
    unique() |> 
    walk(plot_1_variable, x)
}
```


### Running the pipeline

#### All cases

Running the pipeline for all cases vs controls (takes 2'):

```{r eval = FALSE}
output_all_cont <- pipeline()
```

```{r include = FALSE}
if (file_exists("output_all_cont.rds")) {
  output_all_cont <- readRDScache("output_all_cont.rds")
} else {
  output_all_cont <- pipeline()
  saveRDScache(output_all_cont, "output_all_cont.rds")
}
```

Let's compare the p values:

```{r}
output_all_cont |> 
  select(response, timepoint, control, HFMD, starts_with("p_")) |> 
  mutate(across(starts_with("p"), ~ round(.x, 3))) |> 
  print(n = Inf) 
```

Let's look at the estimates:

```{r}
output_all_cont |> 
  select(response, timepoint, control, HFMD, starts_with("cf")) |> 
  mutate(across(starts_with("cf"), ~ round(.x, 2))) |> 
  print(n = Inf) 
```

Computing the diagnostic metrics for all the propensity score weights:

```{r}
c("ow", "tw", "t2") |>
  map(diagnostic_ps, x = output_all_cont) |> 
  walk(print, n = Inf)
```

**Interpretations:**

* covariates balance < 0.1 is good. Values between 0.1 and 0.2 are indicative of
moderate imbalance. Values above 0.2 are a problem.
* weights should not be too high.
* weights distribution should not be too heavy right tail.
* mean an median of weights should not be too different.
* effective sample size should not be too different from real sample size (not less
than 50% of original sample size).
* effective sample size should not be lower than the number treated.
* max weight should not be much higher than 10 to 20 times the median weight.

The number of data points:

```{r}
output_all_cont |> 
  select(response, timepoint, control, HFMD) |> 
  print(n = Inf)
```

Let's plot the effect for all the score, time points and methods:

```{r fig_Bayley, fig.width = 7, fig.height = 25 / 3}
opar <- par(mfrow = c(5, 3), plt = c(.15, .97, .15, .9))
plot_effects_all(output_all_cont)
par(opar)
```


#### Other controls

Severe cases vs controls:

```{r eval = FALSE}
output_sev_cont <- filter_patients(group == "control" |
                                     !(is.na(HFMD) | HFMD == "grade 2a")) |> 
  pipeline()
```

```{r include = FALSE}
if (file_exists("output_sev_cont.rds")) {
  output_sev_cont <- readRDScache("output_sev_cont.rds")
} else {
  output_sev_cont <- filter_patients(group == "control" |
                                       !(is.na(HFMD) | HFMD == "grade 2a")) |> 
  pipeline()
  saveRDScache(output_sev_cont, "output_sev_cont.rds")
}
```

```{r fig_sev_cont, fig.width = 7, fig.height = 25 / 3}
opar <- par(mfrow = c(5, 3), plt = c(.15, .97, .15, .9))
plot_effects_all(output_sev_cont)
par(opar)
```

Non severe cases vs controls:

```{r eval = FALSE}
output_nonsev_cont <- filter_patients(group == "control" |
                                        !(is.na(HFMD) | HFMD != "grade 2a")) |> 
  pipeline()
```

```{r include = FALSE}
if (file_exists("output_nonsev_cont.rds")) {
  output_nonsev_cont <- readRDScache("output_nonsev_cont.rds")
} else {
  output_nonsev_cont <- filter_patients(group == "control" |
                                          !(is.na(HFMD) | HFMD != "grade 2a")) |> 
    pipeline()
  saveRDScache(output_nonsev_cont, "output_nonsev_cont.rds")
}
```

```{r fig_nonsev_cont, fig.width = 7, fig.height = 25 / 3}
opar <- par(mfrow = c(5, 3), plt = c(.15, .97, .15, .9))
plot_effects_all(output_nonsev_cont)
par(opar)
```

Severe vs non-severe (considered as controls):

```{r eval = FALSE}
output_sev_nonsev <- filter_patients(group == "HFMD", !is.na(HFMD)) |> 
  pipeline2(covariab = covariables_sev)
```

```{r include = FALSE}
if (file_exists("output_sev_nonsev.rds")) {
  output_sev_nonsev <- readRDScache("output_sev_nonsev.rds")
} else {
  output_sev_nonsev <- filter_patients(group == "HFMD", !is.na(HFMD)) |> 
    pipeline2(covariab = covariables_sev)
  saveRDScache(output_sev_nonsev, "output_sev_nonsev.rds")
}
```

```{r fig_sev_nonsev, fig.width = 7, fig.height = 25 / 3}
opar <- par(mfrow = c(5, 3), plt = c(.15, .97, .15, .9))
plot_effects_all(output_sev_nonsev)
par(opar)
```

EV-A71 vs controls:

```{r eval = FALSE}
output_EV71_cont <- filter_patients(group == "control" |
                                      !(is.na(PCR) | ! str_detect(PCR, "EV-A71"))) |>
  pipeline()
```

```{r include = FALSE}
if (file_exists("output_EV71_cont.rds")) {
  output_EV71_cont <- readRDScache("output_EV71_cont.rds")
} else {
  output_EV71_cont <- filter_patients(group == "control" |
                                        !(is.na(PCR) | ! str_detect(PCR, "EV-A71"))) |>
    pipeline()
  saveRDScache(output_EV71_cont, "output_EV71_cont.rds")
}
```

```{r fig_EV71_cont, fig.width = 7, fig.height = 25 / 3}
opar <- par(mfrow = c(5, 3), plt = c(.15, .97, .15, .9))
plot_effects_all(output_EV71_cont)
par(opar)
```

Non EV-A71 vs controls:

```{r eval = FALSE}
output_nonEV71_cont <- filter_patients(group == "control" |
                                         !(is.na(PCR) | str_detect(PCR, "EV-A71"))) |>
  pipeline()
```

```{r include = FALSE}
if (file_exists("output_nonEV71_cont.rds")) {
  output_nonEV71_cont <- readRDScache("output_nonEV71_cont.rds")
} else {
  output_nonEV71_cont <- filter_patients(group == "control" |
                                           !(is.na(PCR) | str_detect(PCR, "EV-A71"))) |>
    pipeline()
  saveRDScache(output_nonEV71_cont, "output_nonEV71_cont.rds")
}
```

```{r fig_nonEV71_cont, fig.width = 7, fig.height = 25 / 3}
opar <- par(mfrow = c(5, 3), plt = c(.15, .97, .15, .9))
plot_effects_all(output_nonEV71_cont)
par(opar)
```

EV-A71 vs non-EV-A71 (considered as controls):

```{r eval = FALSE}
output_EV71_nonEV71 <- filter_patients(group == "HFMD", !is.na(PCR)) |> 
  pipeline2(covariab = covariables_EV71)
```

```{r include = FALSE}
if (file_exists("output_EV71_nonEV71.rds")) {
  output_EV71_nonEV71 <- readRDScache("output_EV71_nonEV71.rds")
} else {
  output_EV71_nonEV71 <- filter_patients(group == "HFMD", !is.na(PCR)) |> 
    pipeline2(covariab = covariables_EV71)
  saveRDScache(output_EV71_nonEV71, "output_EV71_nonEV71.rds")
}
```

```{r fig_EV71_nonEV71, fig.width = 7, fig.height = 25 / 3}
opar <- par(mfrow = c(5, 3), plt = c(.15, .97, .15, .9))
plot_effects_all(output_EV71_nonEV71)
par(opar)
```

Lesions on MRI vs controls:

```{r eval = FALSE}
output_lesions_cont <- filter_patients(group == "control" |
                                         ! (is.na(Final) | ! Final)) |>
  pipeline()
```

```{r include = FALSE}
if (file_exists("output_lesions_cont.rds")) {
  output_lesions_cont <- readRDScache("output_lesions_cont.rds")
} else {
  output_lesions_cont <- filter_patients(group == "control" |
                                           ! (is.na(Final) | ! Final)) |>
    pipeline()
  saveRDScache(output_lesions_cont, "output_lesions_cont.rds")
}
```

```{r fig_lesions_cont, fig.width = 7, fig.height = 25 / 3}
opar <- par(mfrow = c(5, 3), plt = c(.15, .97, .15, .9))
plot_effects_all(output_lesions_cont)
par(opar)
```

No lesions on MRI vs controls:

```{r eval = FALSE}
output_nolesions_cont <- filter_patients(group == "control" |
                                         ! (is.na(Final) | Final)) |>
  pipeline()
```

```{r include = FALSE}
if (file_exists("output_nolesions_cont.rds")) {
  output_nolesions_cont <- readRDScache("output_nolesions_cont.rds")
} else {
  output_nolesions_cont <- filter_patients(group == "control" |
                                             ! (is.na(Final) | Final)) |>
    pipeline()
  saveRDScache(output_nolesions_cont, "output_nolesions_cont.rds")
}
```

```{r fig_nolesions_cont, fig.width = 7, fig.height = 25 / 3}
opar <- par(mfrow = c(5, 3), plt = c(.15, .97, .15, .9))
plot_effects_all(output_nolesions_cont)
par(opar)
```

MRI lesions vs no MRI lesions (considered as controls):

```{r}
#output_lesions_nolesions <- filter_patients(group == "HFMD", !is.na(Final)) |> 
#  pipeline2(covariab = covariables_MRI)
```



## Longitudinal analysis

A function that summarizes the number of patients available for various longitudinal
schemes:

```{r}
longitudinal_summary <- function(x) {
  x <- x |>
    group_by(ParNo, time_point) |> 
    tally() |> 
    na.exclude() |> 
    pivot_wider(names_from = time_point, values_from = n)
  out <- c("3 time points" = nrow(na.exclude(x)),
           "at least first 2 time points" = nrow(na.exclude(select(x, -`18 months`))),
           "at least last 2 time points" = nrow(na.exclude(select(x, -enrollment))),
           "at least the first and last time points" =
             nrow(na.exclude(select(x, - `6 months`))))
  tibble(condition            = names(out), 
         "number of patients" = out)
}
```

A function that runs `longitudinal_summary()` on all the scores of a follow-up data
frame `df`:

```{r}
lsummary <- function(df) {
  scores |> 
    set_names() |>
    map(~ longitudinal_summary(df[! is.na(df[[.x]]), ])) |> 
    bind_rows(.id = "score")
}
```

The patients ID of the HFMD cases:

```{r}
HFMDcases <- children |> 
  filter(group == "HFMD") |> 
  pull(ParNo)
```

Looking that the number of case and control patients for which we have longitudinal
data:

```{r}
left_join(rename(lsummary(filter(followups, ParNo %in% HFMDcases)),
                 HFMD = `number of patients`),
          rename(lsummary(filter_out(followups, ParNo %in% HFMDcases)),
                 control = `number of patients`), c("score", "condition")) |> 
  filter(HFMD > 5) |> 
  print(n = Inf)
```

```{r}
hash <- c(enrollment = 1, `6 months` = 2, `18 months` = 3)

followups2 <- followups |> 
  mutate(x = hash[time_point]) |> 
  left_join(select(children, ParNo, group), "ParNo")

ids <- followups2 |> 
  filter_out(is.na(CS)) |> 
  group_by(ParNo) |> 
  tally() |> 
  filter(n > 2) |> 
  pull(ParNo)

followups3 <- followups2 |> 
  filter_out(is.na(CS)) |> 
  select(ParNo, time_point, x, CS, group) |> 
  filter(ParNo %in% ids) |> 
  arrange(ParNo, time_point)

with(followups3, plot(x, CS))

followups3 |> 
  mutate(col = c(HFMD = "red", control = "blue")[group]) |> 
  group_by(ParNo) |> 
  group_split() |> 
  sample() |> 
  walk(~ with(.x, lines(x, CS, col = col)))

fu4 <- followups3 |>
  pivot_wider(id_cols = -x, names_from = time_point, values_from = CS) |> 
  mutate(col = c(HFMD = "red", control = "blue")[group]) |> 
  select(-c(ParNo, group))

fu4 |> 
  select(-`18 months`) |> 
  group_by(across(everything())) |> 
  summarise(n = sqrt(n())) |> 
  with(plot(enrollment, `6 months`, cex = n, col = col))
```




